Explorations to Optimize the Parareal Algorithm \\to Solve ODEs in Given Applications

### 2014-12-10 paper.tex
another hour push

The parareal algorithm, developed by Lions, Maday and Turinici in 2001, CITE is
a generalalized algorithm which allows for parallelization in time. It does so by using a
cheaper ($g_{\Delta t}$)---lower order or lesser resolution, approximation first which it then corrects by evaluating segments with a finer method ($g_{fine}$). The correction step from the approximation can be done in parallel. Multiple iterations can be performed to achieve an error provably asymptotically equal to a full serial computation of $g_{fine}$. The exact number of iterations required to achieve certain levels of convergence depends on the problem and must be tuned top optimize for speedup for a given system. In summary, the algorithm is $k$ repetitions of a
correction process which requires running updated by a finer method run in parallel for subsections.

@@

The beauty is the parareal method can be adapted into many different versions with different choices of the $g_{\Delta t}$ and $g_{fine}$ solution operators.  The $g_{fine}$ solution is to a more computationally expensive method but assumed to provide either greater resolution or lower error.  Most of the time this will be a higher order method, but can also be a greater sampling of the same order method used in the coarse version.  The exact optimum pairing is problem-specific and the code is written so that a different "n-order method step" function could be written and then replaced into the code.

\subsection{Visualization of Scheme}

@@

The first t seconds comes from the first coarse approximation, without which the parareal algorithm degenerates to $g_{\Delta t}$,  In order for there to be a speedup relative to the fine operator, we require
that $t + k(t + Qt/N) < Qt$, or $k < \frac{Q - 1}{1 + Q/N}$ or $N > \frac{Qk}{Q
- 1 - k}$. This can be tuned with by either adjusting the the quality factor or reducing the number of iterations, but the number of iterations, k, is important for error convergence to that of $g_{fine}$.  However, this optimal k is a problem-specific parameter and must be optimized for.

@@

\subsection{Expectations}
TODO WESLEY
\subsection{General Remarks about Resuts}
One of the TODO WESEY

\subsection{Comparison to Serial with $g_{fine}$ as a Forward Euler with
Smaller $\Delta t$}

@@

Varying Iterations:
]
Varying Quality Factor:
We define the quality factor in this case for how many times smaller the Euler method used for $g_{fine}$ is compared to the Euler method used for $g_{\Delta t}$

@@

\subsection{Comparison to Serial with $g_{fine}$ as Higher Order Methods}

@@

\subsection{Strong and Weak Scaling Analysis}

ashi TODO by taking some numbers and following
\url{https://www.sharcnet.ca/help/index.php/Measuring_Parallel_Scaling_Performance}

@@

@@

\subsection{Possible Optimizations}

@@

>>>>

The parareal algorithm, developed by Lions, Maday and Turinici in 2001, CITE is
a generalalized algorithm which allows for parallelization in time. It does so by using a
cheaper ($g_{\Delta t}$)---lower order or lesser resolution, approximation first which it then corrects by evaluating segments with a finer method ($g_{\textnormal{fine}}$). The correction step from the approximation can be done in parallel. Multiple iterations can be performed to achieve an error provably asymptotically equal to a full serial computation of $g_{\textnormal{fine}}$. The exact number of iterations required to achieve certain levels of convergence depends on the problem and must be tuned top optimize for speedup for a given system. In summary, the algorithm is $k$ repetitions of a
correction process which requires running updated by a finer method run in parallel for subsections.

@@

The beauty of the parareal method is that it can be adapted into many different versions with different choices of the $g_{\Delta t}$ and $g_{\textnormal{fine}}$ solution operators. The $g_{\textnormal{fine}}$ solution is a more computationally expensive method but assumed to have lesser error.  Most of the time this will be a higher order method, but can also be a greater sampling of the same order method used in the coarse version.  The exact optimum pairing is problem-specific and the code is written so that a different "n-order method step" function could be written and then replaced into the code. We experiment with testing different pairs of $g_{\Delta t}$ and $g_{\textnormal{fine}}$ to see how influential the choice of solution scheme is.

\subsection{Visualization of Parareal Scheme}

@@

The first t seconds comes from the first coarse approximation, without which the parareal algorithm degenerates to $g_{\Delta t}$,  In order for there to be a speedup relative to the fine operator, we require
that $t + k(t + Qt/N) < Qt$, or $k < \frac{Q - 1}{1 + Q/N}$ or $N > \frac{Qk}{Q
- 1 - k}$. This can be tuned with by either adjusting the the quality factor or reducing the number of iterations, but the number of iterations, k, is important for error convergence to that of $g_{\textnormal{fine}}$.  However, this optimal k is a problem-specific parameter and must be optimized for.

TODO PARALLEL IN SPACE ANALYSIS

\subsection{Incorporating Communication}

TODO PARALLEL IN SPACE ANALYSIS

\subsection{Amdahl's Law Analysis}

TODO PARALLEL IN SPACE ANALYSIS

@@

\subsection{Expectations}
TODO WESLEY
\subsection{General Remarks about Resuts}
One of the TODO WESLEY

\subsection{Comparison to Serial with $g_{\textnormal{fine}}$ as a Forward Euler with
Smaller $\Delta t$}

@@

Varying Iterations:
]
Varying Quality Factor:
We define the quality factor in this case for how many times smaller the Euler method used for $g_{\textnormal{fine}}$ is compared to the Euler method used for $g_{\Delta t}$

@@

\subsection{Comparison to Serial with $g_{\textnormal{fine}}$ as Higher Order Methods}

@@

@@

ashi TODO by taking some numbers and following

@@

\subsection{Possible Optimizations and Future Work}

@@

TODO Wesley
